---
title: Selling Weekly Strangles on SPY
author: "Pritam Dalal"
date: "11/28/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include=FALSE}
######################
## loading packages ##
######################
library(tidyverse)
df_pnl <- read_csv("../data_output/spy_weekly_pnl.csv", col_types = cols())
getwd()
```

##### Backtest Parameters
Selling weekly strangles on SPY naked, no delta-hedging.  Accounts for bid-ask spread, but not commissions.  Tested 50-delta, 30-delta, 10-delta.  At the moment, the three variations are not equally sized, so comparing total PNL is not meaninful.  Sharpe-Ratio is meaningful.

Dates: 12/16/2018 - 7/27/2018.

##### TTD PNL

At the end of the day, all three were losers, largely because of massive drawdowns at the beginning of 2018.

```{r, ttd_pnl, echo = FALSE}
# cummulative PNL graph
df_pnl %>% 
    group_by(variation, data_date) %>% 
    summarize(
        daily = sum(dly_opt_pnl)
    ) %>% 
    mutate(
        ttd = cumsum(daily)
    ) %>% 
    ggplot() +
        geom_line(aes(x = data_date, y = ttd, color = factor(variation)))
 
```

## Sharpe-Ratio During 2017

Trading options is good...until it isn't.

```{r pressure, echo=TRUE}
# sharpe in 2017
df_pnl %>% 
    dplyr::filter(data_date < "2018-01-01") %>% 
    group_by(variation) %>% 
    summarize(
        sharpe_ratio = (mean(dly_opt_pnl) / sd(dly_opt_pnl)) * sqrt(252)
    )
```

