---
title: "Naked Puts vs Delta-Hedged Puts"
author: "Pritam Dalal"
output: 
  html_document:
    theme: simplex
    df_print: kable
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, load_packages, include = FALSE}
# loading packages
library(tidyverse)
library(tidyquant)
```

```{r, read_data, include = FALSE}
###################
# reading in data #
###################
df_underlying_all <-
    read_csv("../data_input/monthly_underlying.csv", col_type = cols())
df_monthly_pnl_all <- 
    read_csv("../../delta_neutral_data_output/monthly_strangle_pnl.csv",col_type = cols())
df_spy_weekly_pnl_all <- 
    read_csv("../../delta_neutral_data_output/spy_weekly_strangle_pnl.csv", col_type = cols())
df_monthly_position_size_all <- 
    read_csv("../../delta_neutral_data_output/monthly_position_scaling.csv", col_type = cols())
df_spy_weekly_position_size_all <- 
    read_csv("../../delta_neutral_data_output/spy_weekly_position_scaling.csv", col_type = cols())

dbl_aum <- 100000
dbl_base_size <- 75
dbl_aum_mult <- dbl_aum / dbl_base_size
```


```{r, combine_dataframes, include = FALSE}
####################################
## combining dataframes as needed ##
####################################
# the purpose of this code is to combine the pnls from the weekly strategy and the monthly strategy into one dataframe so that we can efficiently calculate statistics on them. 

# adding frequency columns to pnl dataframes
df_monthly_pnl_all <- 
    df_monthly_pnl_all %>% mutate(frequency = "monthly")
df_spy_weekly_pnl_all <- 
    df_spy_weekly_pnl_all %>% mutate(frequency = "weekly")

# combing pnl dataframes
df_pnl_all <- bind_rows(df_monthly_pnl_all, df_spy_weekly_pnl_all)

# adding frequency columns to position size dataframes
df_monthly_position_size_all <-
    df_monthly_position_size_all %>% mutate(frequency = "monthly")
df_spy_weekly_position_size_all <- 
    df_spy_weekly_position_size_all %>% mutate(frequency = "weekly")

# combining position size dataframes
df_position_size_all <- 
    bind_rows(df_monthly_position_size_all, df_spy_weekly_position_size_all)
```




```{r, selecting_underlyings, include = FALSE}
df_underlying <-
    df_underlying_all %>% 
    filter(underlying == "SPY")
```


```{r, filtering_data, include = FALSE}
############################################
## filtering and adding columns as needed ##
############################################
# position size - filtering by underlying and variation
df_position_size <- 
    df_position_size_all %>% 
    filter(underlying %in% df_underlying$underlying) %>%
    filter(variation == 0.1)

    
# pnls - scaling the pnl and rearranging columns
df_pnl <-         
    df_pnl_all %>% 
    filter(underlying_symbol %in% df_underlying$underlying) %>%
    filter(variation == 0.1) %>% 
    mutate(
        naked_dly  = dly_opt_pnl * strangle_mult * dbl_aum_mult
        , dh_dly = dly_tot_pnl * strangle_mult * dbl_aum_mult
    ) %>% 
    select(frequency, underlying_symbol:strangle_mult, naked_dly:dh_dly) %>% 
    arrange(frequency, underlying_symbol, type, expiration, data_date)
```

```{r, spy_performance, include = FALSE}
# getting SPY data to throw in the comparison, using tidyquant package
df_spy_px <- 
    tq_get("SPY", get = "stock.prices", from = "2016-12-16", to = "2018-07-23")
# calculating log returns
df_spy_px$ret <- 
    log(df_spy_px$adjusted) - log(lag(df_spy_px$adjusted, 1)) 
# calculating PNL
df_spy_px <-
    df_spy_px %>% 
        mutate(
            pnl = dbl_aum * ret
        ) 
```


&nbsp;

The purpose of this analysis is to demonstrate the capabilities Option Data Science (ODS) backtesting framework.  In order to do this we examine four different SPY option premium selling strategies.  Specifically, we consider the repeated selling of weekly and monthly options, both naked and delta-hedged.

**NOTHING**

&nbsp;

#### Backtesting Parameters

**Dates:** 12/16/2016 - 7/20/2018 (400 total trading days)

**Underlying:** SPY

**AUM:** $100,000

**Weekly Strategies:** on each Friday, sell 10-delta strangles that expire the following Friday.

**Monthly Strategies:** on each regular option expiration (3rd Friday of the month), sell 10-delta strangles that expire on the next regular expiration.


For both strategies the firest trades open on 12/16/2018 and the final trades close on 7/20/2018.  The monthly strategy consists is executed over 19 consecutive regular expirations, while the weekly strategy is executed over 83 consectutive weekly expirations.  We will examine naked and delta-hedged variants of both strategies.


Let's consider the total the amount of premium sold, and number of options sold, for each strategy.  Here is a table of summary statistics: 
```{r, und_exp_prem_sold, echo = FALSE}
df_position_size %>% 
    group_by(frequency) %>% 
    summarize(
        tot_prem = sum(strangle_prem_sold) * dbl_aum_mult
        , avg_prem = mean(strangle_prem_sold) * dbl_aum_mult
        , tot_opt = sum(strangle_mult) * dbl_aum_mult / 100
        , avg_opt = mean(strangle_mult) * dbl_aum_mult / 100
        #, min_prem = min(strangle_prem_sold) * dbl_aum_mult
        #, max_prem = max(strangle_prem_sold) * dbl_aum_mult
    ) %>% 
    .[c(2, 1), ]
```

Notice that both strategies sell about \$26K in premium over the backtest period.  The weekly strategies sell about \$300 in premium per execution, while the monthly strategies sell about \$1,375 in premium per expiration.  Over the entire backtest period, the weekly strategies sell 858 options, which is 2.5 times as many as the monthly strategies.  This makes sense because the options being sold in the weekly strategy are of a much shorter tenor, and are therefore worth quite a bit less.


&nbsp;

#### Performance
Let's compare the performance of the four strategies:
```{r}
########################################################
## gathered the PNLs together and performing analysis ##
########################################################
# daily pnl for each strategy, all underlyings combined
df_dly_pnl <-
    df_pnl %>%
        group_by(frequency, data_date) %>%
        arrange(frequency, data_date) %>%
        summarize(
            naked_dly = sum(naked_dly)
            , dh_dly = sum(dh_dly)
        )

# gathering the dh_pnl column
df_pnl_gathered <-  
    df_dly_pnl %>% 
        rename(naked = naked_dly, dh = dh_dly) %>% 
        gather(hedge, daily_pnl, naked:dh) %>% 
        mutate(strategy = paste0(frequency, "_", hedge)) %>% 
        ungroup() %>% 
        select(strategy, data_date, daily_pnl) %>% 
        bind_rows(
            df_spy_px %>% 
                mutate(
                    strategy = "spy"
                    , daily_pnl = replace_na(pnl, 0)
                ) %>%
                select(strategy, data_date = date, daily_pnl)
        ) 


# calculating trade-to-date PNLs
df_pnl_gathered <- 
    df_pnl_gathered %>% 
        group_by(strategy) %>% 
        mutate(
            ttd_pnl = with_order(data_date, cumsum, daily_pnl)
        ) %>% 
        mutate(
            draw_down = cummax(ttd_pnl) - ttd_pnl
        )


# calculating various performance metrics
df_performance <- 
    df_pnl_gathered %>% 
        filter(strategy != "spy") %>% 
        group_by(strategy) %>% 
        summarize(
            tot_pnl = sum(daily_pnl) %>% round(2)
            , sharpe_ratio = 
                ((mean(daily_pnl) / sd(daily_pnl)) * sqrt(252) )%>% round(2)
            , ret = 
                ((((sum(daily_pnl) + dbl_aum) / dbl_aum) ^ (252/400)) - 1) %>% 
                round(3)
            , max_drawdown = (max(draw_down) / dbl_aum) %>% round(2)
        ) %>% 
        .[c(3, 4, 1, 2), ]

```




```{r, calc_performance, include = FALSE}
##############
## OLD CODE ##
##############
# 
# # daily pnl for each strategy, all underlyings combined
# df_dly_pnl <-
#     df_pnl %>% 
#         group_by(frequency, data_date) %>%
#         arrange(frequency, data_date) %>% 
#         summarize(
#             naked_dly = sum(naked_dly)
#             , dh_dly = sum(dh_dly) 
#         )
# 
# 
# # total PNL for both strategies
# tot_unhedged <- df_dly_pnl$naked_dly %>% sum() %>% round(2)
# tot_hedged <- df_dly_pnl$dh_dly %>% sum() %>% round(2)
# tot_spy <- df_spy_px$pnl %>% sum(na.rm = TRUE) %>% round(2)
# 
# 
# # mean
# avg_unhedged <- df_dly_pnl$naked_dly %>% mean() %>% round(2)
# avg_hedged <- df_dly_pnl$dh_dly %>% mean() %>% round(2)
# avg_spy <- df_spy_px$pnl %>% mean(na.rm = TRUE) %>% round(2)
# 
# # standard deviation
# sd_unhedged <- df_dly_pnl$naked_dly %>% sd() %>% round(2)
# sd_hedged <- df_dly_pnl$dh_dly %>% sd() %>% round(2)
# sd_spy <- df_spy_px$pnl %>% sd(na.rm = TRUE) %>% round(2)
# 
# 
# # sharpe-ratio
# sr_unhedged <- 
#     (mean(df_dly_pnl$naked_dly) / sd(df_dly_pnl$naked_dly)) * sqrt(252) %>% round(2) # naked
# sr_hedged <- 
#     (mean(df_dly_pnl$dh_dly) / sd(df_dly_pnl$dh_dly)) * sqrt(252)  %>% round(2)# hedged
# sr_spy <- 
# (mean(df_spy_px$pnl, na.rm = TRUE) /
#         sd(df_spy_px$pnl, na.rm = TRUE)) * sqrt(252)  %>% round(2) # SPY
# 
# # returns
# ret_unhedged <- (101.27/100)^(365/581) - 1 # naked
# ret_hedged <- (103.22/100)^(365/581) - 1 # delta-hedged
# ret_spy <- (df_spy_px$adjusted[nrow(df_spy_px)]
#     /df_spy_px$adjusted[1]) ^ (365/581) - 1 # SPY 
# 
# # DVAR
# dvar_hedged <- quantile(df_dly_pnl$naked_dly, 0.05) %>% round(2)
# dvar_unhedged <- quantile(df_dly_pnl$dh_dly, 0.05) %>% round(2)
# dvar_spy <- quantile(df_spy_px$pnl, 0.05, na.rm = TRUE) %>% round(2)
# 
# df_performance <-
#     tibble(
#         strategy = c("unhedged", "hedged", "spy")
#         , total = c(tot_unhedged, tot_hedged, tot_spy)
#         , mean = c(avg_unhedged, avg_hedged, avg_spy)
#         , sd = c(sd_unhedged, sd_hedged, sd_spy)
#         , sharpe = c(sr_unhedged, sr_hedged, sr_spy)
#         , ret = c(ret_unhedged, ret_hedged, ret_spy)
#         , dvar = c(dvar_hedged, dvar_unhedged, dvar_spy)
#     )
```



```{r, performance_table, echo = FALSE}
df_performance
```

None of these strategies performed particularly well during the backtest period, especially considering that SPY had 15\% a return and a 1.45 Sharpe-Ratio during the same timeframe.  All strategies suffered significant drawdowns during market turmoil in early 2018; the monthly-naked strategy lost a staggering 30\% of AUM before rebounding.

&nbsp;

#### Visualizing Cummulative Performance

As we can see from the table above, delta-hedging reduces the risk significantly, but the risk is not zero. Graphing the cummulative PNLs of the hedged and unhedged strategies is another way to illustrate this risk reduction:
```{r, ttd_pnl_graph, echo = FALSE}
# df_pnl %>% 
#     group_by(data_date) %>%
#     arrange(data_date) %>% 
#     summarize(
#         naked_dly = sum(naked_dly)
#         , dh_dly = sum(dh_dly) 
#     ) %>% 
#     mutate(
#         naked_ttd = cumsum(naked_dly)
#         , dh_ttd = cumsum(dh_dly)
#     ) %>% 
#     ggplot() +
#         geom_line(aes(x = data_date, y = naked_ttd), color = "green") +
#         geom_line(aes(x = data_date, y = dh_ttd), color = "blue")

df_pnl_gathered %>% 
    filter(strategy != "spy") %>% 
    ggplot() +
    geom_line(aes(x = data_date, y = ttd_pnl, color = strategy))

```

The graph of the unhedged strategy is instructive of the nature of premium selling.  Everthing is great...until it isn't.

&nbsp;

#### How to Improve This Analysis

1. Add labeling to all graphs.

1. Include SPY in ttd_pnl graph.

1. Add legend to PNL graph.

1. Additional performance metrics: largest draw down.

1. Graph PNLs by month - this is important in the alternative investment space.

1. Make max width of the writing more narrow.

1. Title font is too big.

1. Change table column names to be more descriptive.

1. Change returns to percentage,







